---
title: "OHI-atl/npp prs_fish layer/npp"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(dplyr)        #NOT spatial - this is a data wrangling library
library(ggplot2)
library(here)
library(tidyverse)
library(ncdf4)
library(raster)
library(RColorBrewer)
library(sf)
```

#Summary
--> to do: (from ohiprep_v2020 and previous) This script calculates log-transformed, mean annual net primary production values globally from 2003 to 2015. This data is used to standardize commercial fishing catch to produce [fishing pressure layers for OHI]

#Data source
GLOBAL-REANALYSIS-BIO-001-029-MONTHLY

**Reference**:CMEMS copernicus https://resources.marine.copernicus.eu/?option=com_csw&task=results **Downloaded**: August 12, 2020 Description: Monthly Net Primary Production (mg C / m2 / day): we used Total Primary Production of Phyto
nppv (mg m-3 day-1) 
**Native data** resolution: 0.25 degree x 0.25 degree (75 depth levels) 
**Time range**: From 1993-01-01 to 2019-12-23. both daily and montly mean. 
**Format**: NetCfd fo rmat Variables: CHL PHYC O2 NO3 PO4 SI FE SPCO2 PH PP. Here we used PP. **Geographical area selected**: -99, 85, 70, -80

IL FILE è TROPPO GRANDE. DEVO FORSE SCARICARLI IN CARTELLE ANNO PER ANNO PER POTERCI FARE LA MEDIA PER CIASCUN ANNO? Instructions

Parameters used for downloading the data

Parameter	Value|Product	MEDSEA_ANALYSIS_FORECAST_BIO_006_014
---------------|--------------------------------------------
Frequency|	monthly
Lat min|	30.18
Lat max|	45.97
Lon min|	-5.54
Lon max|	36.29
Timesteps|	2019-09-01, 2019-09-09
Service for downloading|	
Files total dimension|	~250 MB
Datasets| 
* med-ogs-pft-an-fc-m
* med-ogs-nut-an-fc-m
* med-ogs-co2-an-fc-m

```{r}
#import shp for the regions
regions_atl_FAO <- st_read("~/big/atl_regions/Regions_atl_FAO.shp", quiet = TRUE)
reg_atl_FAO_sp <- as(regions_atl_FAO, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)
reg_atl_FAO_sp
plot(reg_atl_FAO_sp)
      

      #import shp for the regions
atl_mask025_vector <- st_read("~/big/prep/spatial/atl_mask025_vector.shp", quiet = TRUE)
atl_mask025_vector_sp <- as(atl_mask025_vector, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)
reg_atl_FAO_sp
plot(atl_mask025_vector_sp)

# set bigdir directory for raster analysis
setwd("C:/Users/Acer/Documents/big")
R_raster_tmp='~/big/R_raster_tmp' 
dir.create(R_raster_tmp, showWarnings=T)
dir_B='~/big' 
#define paths for the raw data and layer folder held on my pc Documents folder. (Once the project starts, the path should migrate to a shared database)

npp_19 <- stack(file.path(R_raster_tmp, 'global-reanalysis-bio-001-029-monthly_1597267552177.nc'))#Large SpatialGridDataFrame

#You can perform operations on a RasterStack by using the calc() function from the raster package. calc() lets you define a function to apply across all layers in the stack.
atl_mask025<-raster::raster(file.path(dir_B,"prep/spatial/atl_mask_final_cmems025.tif"))
plot(atl_mask025)
nppAvg = raster::calc(npp_19, fun=function(x){mean(x, na.rm=T)})
nppAvg_c<-crop(nppAvg,atl_mask025_vector_sp)
  plot(nppAvg_c)
nppAvg_m<-mask(nppAvg_c,atl_mask025_vector_sp)
crs(atl_mask025)=crs(nppAvg)
plot(nppAvg_m)
```


```{r}
### Define spectral color scheme for plotting maps
 cols = rev(colorRampPalette(brewer.pal(9, 'Spectral'))(255)) # rainbow color scheme 

plot(nppAvg_m, main = 'Mean Total Primary Production of Phyto (mg m-3 day-1)', col=cols, axes=FALSE)+
    plot(atl_mask025_vector_sp, add=TRUE,axes=FALSE)

##turning the reg shapefile into a raster with the same dimensions as npp_crop but with cell values equal to the rgn_id of each polygon in the shapefile

    ## By adding 'filename=......tif' R will directly save the raster into the defined file rather than memory
zones = rasterize(reg_atl_FAO_sp, nppAvg_m, field = reg_atl_FAO_sp@data$ID )#, filename='~/big/atl_regions/zones.tif')
 {plot(zones, axes=F,main = "Zonal Raster for Atlantic Ocean \n (cell values are unique to sub-regions)")+
  plot(nppAvg_m, add=TRUE,col=cols)+
  plot(reg_atl_FAO_sp, add=TRUE)}#{}+ are necessary for rmd.
mean_z <- zonal(nppAvg_m,zones,fun=mean) #we want to get the mean per region, this could also be min/max or another function
mean_z
plot(mean_z)

grid_1_cmems_atl <- st_read("~/big/prep/spatial/qgis/grid_1_cmems_atl.shp", quiet = TRUE)
grid_1_cmems_atl_sp <- as(grid_1_cmems_atl, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)

plot(grid_1_cmems_atl_sp)
```


#`extract()`

Use `raster::extract()` to get all values of the raster within each of your subregions. This function returns all cell alues in a list in length equal to the number of subregions. The list can be turned into a dataframe with two columns;

- **value** are the cell values (in this case number of anomalous weeks)  
- **rgn_name** are the names of the sub regions

```{r}
# get all values within each region


system.time(extr <- raster::extract(nppAvg_m,grid_1_cmems_atl_sp,method='simple',weight=TRUE)) #weight:true->how much of the cell falls into each region

%>%
  setNames(reg_atl_FAO_sp@data$NAME_EN))
# plot distribution of data per region

df <- data.frame(unlist(extr))%>%
      rename(value = unlist.extr.)%>%  
        filter(!is.na(value)) %>% 
      mutate(id=row_number())

#the gsub here removes all numbers from the rgn_name
#now we have each value assigned to rgn_name.
head(df)
plot(df)
plot(extr[,"value"])

ggplot(df)
```


```{r}
vals = raster::extract(nppAvg_c,reg_atl_FAO_sp,method='simple')%>%
  setNames(reg_atl_FAO_sp@data$NAME_EN)
# plot distribution of data per region
df <- data.frame(unlist(vals))%>%
  rename(value = unlist.vals.)%>%
  mutate(rgn_name = gsub("\\d+", "",row.names(.))) #the gsub here removes all numbers from the rgn_name
#now we have each value assigned to rgn_name.
head(df)


mean_df <- df%>%
  group_by(rgn_name)%>%
  summarise(mean_nnp = mean(value,na.rm=T))
mean_df


print(raw)
#Crop raster data to region extent



```
