---
title: "Regions -Select EEZs from Global Shapefile and FAO shp"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(dplyr)        #NOT spatial - this is a data wrangling library

#devtools::install_github("ecohealthalliance/fasterize")  #to install fasterize you need to run this line
library(fasterize)
#devtools::install_github("tidyverse/ggplot2")
library(ggplot2)
library(dplyr)
library(rgdal)
library(raster)
library(here)
library(ggridges)
library(ggplot2)
library(tidyverse)
library(ncdf4)

```


#Read in global shapefile

You will have to point the path to where the global OHI EEZ file is located on your local computer.

#use st_read to read in the global regions shapefile as a simple feature object -->https://ohi-science.org/news/cropping-rasters-down-to-size 
This comes from the sf package and is a preferred method of loading and working with spatial data as opposed to using readOGR() from the rgdal package. The st_read() function will read in the shapefile as a simple feature object, meaning the spatial information is stored in a dataframe. This allows you to use data wrangling tools such as dplyr to work with the data.
```{r shapefile}
here()

setwd("C:/Users/Acer/Documents/big")
regions_gl <- st_read("~/big/FAO_areasMAJOR/FAO_AREAS.shp", quiet = TRUE)

head(regions_gl)
plot(regions_gl[4]) #plot() which column to show, otherwise R will create a plot for every column, which we do not need

# filter the regions you want:
regions_atl_FAO<- regions_gl %>% 
  dplyr::filter(OCEAN == "Atlantic") %>% 
  filter(F_LEVEL=="MAJOR") %>% 
  select(-13,-14) 
  
#reduce to F_CODE
plot(regions_atl_FAO[4], key.pos=1)
```
```{r Spatial polygon dataframe}
 
#Save shapefile: save the output (saving as an ESRI Shapefile)

setwd("C:/Users/Acer/Documents/big")
bigdir='~/big/R_raster_tmp' 

dir.create(bigdir, showWarnings=T)
sf::write_sf(regions_atl_FAO,paste0(bigdir),"/",  "Regions_atl_FAO.shp", driver = "ESRI Shapefile")#if you want to re-run, delete the file in the folder

#we need to convert the regions object to a SpatialPolygonsDataFrame. This is done using as(). You pass your simple feature object - regions - to as() and tell it to convert to a Spatial object. This returns the same polygons but as a SpatialPolygonsDataFrame.

library(rgdal)
reg_atl_FAO_sp <- as(regions_atl_FAO, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)
reg_atl_FAO_sp

plot(reg_atl_FAO_sp)
```
```{r}

#raster of the atlantic FAO regions created with zonal in npp_pres layer
zones <- stack(file.path(bigdir, 'zones.tif'))#Large SpatialGridDataFrame


library(ncdfgeom)
library(maptools)
data(wrld_simpl)
plot(wrld_simpl)

```

