---
title: "Regions -Select EEZs from Global Shapefile and FAO shp"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(dplyr)        #NOT spatial - this is a data wrangling library

#devtools::install_github("ecohealthalliance/fasterize")  #to install fasterize you need to run this line
library(fasterize)
#devtools::install_github("tidyverse/ggplot2")
library(ggplot2)
library(dplyr)
library(rgdal)
library(raster)
library(here)
library(ggridges)
library(ggplot2)
library(tidyverse)
library(ncdf4)
library(rgeos)
```


#Read in global shapefile

You will have to point the path to where the global OHI EEZ file is located on your local computer.

#use st_read to read in the global regions shapefile as a simple feature object -->https://ohi-science.org/news/cropping-rasters-down-to-size 
This comes from the sf package and is a preferred method of loading and working with spatial data as opposed to using readOGR() from the rgdal package. The st_read() function will read in the shapefile as a simple feature object, meaning the spatial information is stored in a dataframe. This allows you to use data wrangling tools such as dplyr to work with the data.

#Data source

FAO data : http://www.fao.org/figis/geoserver/area/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=area:FAO_AREAS&outputFormat=SHAPE-ZIP

```{r shapefile}
here()



setwd("C:/Users/Acer/Documents/big")
regions_gl <- st_read("~/big/prep/spatial/atl_regions/FAO_AREAS.shp", quiet = TRUE)

head(regions_gl)
plot(regions_gl[4]) #plot() which column to show, otherwise R will create a plot for every column, which we do not need

# filter the regions you want:
regions_atl_FAO<- regions_gl %>% 
  dplyr::filter(OCEAN == "Atlantic") %>% 
  filter(F_LEVEL=="MAJOR") %>% 
  dplyr::select(-13,-14) 
  
#reduce to F_CODE
plot(regions_atl_FAO[4], key.pos=1)
```
```{r Spatial polygon dataframe}
 
#Save shapefile: save the output (saving as an ESRI Shapefile)

#setwd("C:/Users/Acer/Documents/big")
#bigdir='~/big/atl_regions' 
spatial_output='~/big/prep/spatial/output' 
dir.create(bigdir, showWarnings=T)
sf::write_sf(regions_atl_FAO,paste0(spatial_output),  "Regions_atl_FAO.shp", driver = "ESRI Shapefile")#if you want to re-run, delete the file in the folder

regions_atl_FAO <- st_read("~/big/prep/spatial/output/Regions_atl_FAO.shp", quiet = TRUE)

plot(regions_atl_FAO[1])
#we need to convert the regions object to a SpatialPolygonsDataFrame. This is done using as(). You pass your simple feature object - regions - to as() and tell it to convert to a Spatial object. This returns the same polygons but as a SpatialPolygonsDataFrame.

library(rgdal)
reg_atl_FAO_sp <- as(regions_atl_FAO, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)
reg_atl_FAO_sp


plot(reg_atl_FAO_sp)
```

```{r}

#raster of the atlantic FAO regions created with zonal in npp_pres layer
zones <- stack(file.path(bigdir, 'zones.tif'))#Large SpatialGridDataFrame


library(ncdfgeom)
library(maptools)
data(wrld_simpl)
plot(wrld_simpl)

```
##Longhurst provinces
```{r}
lng_gl<-st_read("~/big/Longhurst_world_v4_2010/Longhurst_world_v4_2010.shp",quiet=TRUE)#I have simplified the geometry through the mapshaper.org. The problem lies in some polygons of the baltic sea.
#
lng_gl_sp <- as(lng_gl, 'Spatial')

#lng_gl_sptr <- spTransform(lng_gl_sp, CRS( "+init=epsg:4326" ) ) 
lng_gl_sp_buffer <- rgeos::gBuffer(lng_gl_sp, byid=TRUE, width=0) # go in to solve topology problems


head(lng_gl)
crs(lng_gl_sp)
plot(lng_gl_sp)
head(lng_gl_sptr)
lng_atl_cr<-raster::crop(lng_gl_sp_buffer,reg_atl_FAO_sp)

plot(lng_atl_cr)
head(lng_atl_cr)



```



#Save shapefile

Use rgdal::writeOGR() to save the shapefile. Set the driver equal to ESRI Shapefile.

```{r save}
writeOGR(lng_atl_cr,dsn='~/big/prep/spatial/output',layer='atl_ohi_longhurst', driver = "ESRI Shapefile")

#save a copy on github
writeOGR(lng_atl_cr,dsn='C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output',layer='reg_atl_LNG', driver = "ESRI Shapefile")

reg_atl_LNG <- st_read("C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output/atl_ohi_longhurst.shp", quiet = TRUE)

plot(reg_atl_LNG)
```


##Longhurst provinces

```{r}

lme_gl<-st_read("~/big/prep/spatial/raw/LMEs66.shp",quiet=TRUE)
#I have simplified the geometry through the mapshaper.org. The problem lies in some polygons of the baltic sea.
#
lme_gl_sp <- as(lme_gl, 'Spatial')

lme_gl_sp_buffer <- rgeos::gBuffer(lme_gl_sp, byid=TRUE, width=0) # go in to solve topology problems


head(lme_gl)
crs(lme_gl_sp)==crs(reg_atl_FAO_sp)
plot(lme_gl_sp)
head(lng_gl_sptr)
lme_atl_cr<-raster::crop(lme_gl_sp_buffer,reg_atl_FAO_sp)

plot(lme_atl_cr)
head(lng_atl_cr)

```

```
```{r}
```{r save}
writeOGR(lme_atl_cr,dsn='~/big/prep/spatial/output',layer='reg_atl_lme', driver = "ESRI Shapefile")

#save a copy on github
writeOGR(lme_atl_cr,dsn='C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output',layer='reg_atl_lme', driver = "ESRI Shapefile")

reg_atl_lme <- st_read("C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output/reg_atl_lme.shp", quiet = TRUE)

plot(reg_atl_lme[3])
```

