---
title: "Regions -Select EEZs from Global Shapefile and FAO shp"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)       #Main raster library with nearly all functions used in this analysis
library(rgdal)        #Spatial library - most functions used from rgdal are for vectors (shapefiles)
library(dplyr)        #NOT spatial - this is a data wrangling library

#devtools::install_github("ecohealthalliance/fasterize")  #to install fasterize you need to run this line
library(fasterize)
#devtools::install_github("tidyverse/ggplot2")
library(ggplot2)
library(dplyr)
library(rgdal)
library(raster)
library(here)
library(ggridges)
library(ggplot2)
library(tidyverse)
library(ncdf4)
library(rgeos)
```

## Set template ocean raster and mollweide projection CRS
```{r}
ocean <- raster::raster(file.path(dir_B, 'prep/spatial/spatial_ohi_supplement/ocean.tif'))
mollCRS=crs('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')
```
#Read in global shapefile

You will have to point the path to where the global OHI EEZ file is located on your local computer.

#use st_read to read in the global regions shapefile as a simple feature object -->https://ohi-science.org/news/cropping-rasters-down-to-size 
This comes from the sf package and is a preferred method of loading and working with spatial data as opposed to using readOGR() from the rgdal package. The st_read() function will read in the shapefile as a simple feature object, meaning the spatial information is stored in a dataframe. This allows you to use data wrangling tools such as dplyr to work with the data.

#Data source

https://mazu.nceas.ucsb.edu/data/#ohi_regions

FAO data : http://www.fao.org/figis/geoserver/area/ows?service=WFS&request=GetFeature&version=1.0.0&typeName=area:FAO_AREAS&outputFormat=SHAPE-ZIP

```{r shapefile}
here()
##Read 
regions_ohi_supp<-st_read("~/big/prep/spatial/spatial_ohi_supplement/regions/regions_gcs.shp") %>% 
  filter(rgn_typ%in%c("eez","fao"))#from 322 to 254


setwd("C:/Users/Acer/Documents/big")
regions_gl <- st_read("~/big/prep/spatial/FAO_AREAS.shp", quiet = TRUE)

head(regions_gl)
plot(regions_gl[4]) #plot() which column to show, otherwise R will create a plot for every column, which we do not need

# filter the regions you want:
regions_atl_FAO<- regions_gl %>% 
  dplyr::filter(OCEAN == "Atlantic") %>% 
  filter(F_LEVEL=="MAJOR") %>% 
  dplyr::select(-13,-14) 
  
#reduce to F_CODE
plot(regions_atl_FAO[4], key.pos=1)

######Atlantic mask raster#####
#fasterize the atl_fao to have a mask to crop next file
atl_mask<-fasterize(regions_atl_FAO, ocean, field= "F_STATUS")

##keep cells_raster from Watson code file to use as raster model to fasterizze.
plot(cells_raster)
atl_mask<-fasterize(regions_atl_FAO, cells_raster, field= "F_STATUS")
plot(atl_mask)
writeRaster(atl_mask,
            file.path(dir_B, 'prep/spatial/output/atl_mask.tif'))

atl_mask<-raster::raster("C:/Users/Acer/Documents/big/prep/spatial/atl_mask.tif")
extent(atl_mask)
res(atl_mask)
crs(atl_mask)
plot(atl_mask)
```

```{r Spatial polygon dataframe}
 
#Save shapefile: save the output (saving as an ESRI Shapefile)


spatial_output='~/big/prep/spatial/output' 
dir.create(bigdir, showWarnings=T)
#sf::write_sf(regions_atl_FAO,paste0(spatial_output),  "Regions_atl_FAO.shp", driver = "ESRI Shapefile")#if you want to re-run, delete the file in the folder

regions_atl_FAO <- st_read("~/big/prep/spatial/output/Regions_atl_FAO.shp", quiet = TRUE)

plot(regions_atl_FAO[1])
#we need to convert the regions object to a SpatialPolygonsDataFrame. This is done using as(). You pass your simple feature object - regions - to as() and tell it to convert to a Spatial object. This returns the same polygons but as a SpatialPolygonsDataFrame.

library(rgdal)
reg_atl_FAO_sp <- as(regions_atl_FAO, 'Spatial') #to avoid the error 'invalid multibyte string' delete the special chr (french and spanish name)
reg_atl_FAO_sp


plot(reg_atl_FAO_sp)

rgns_atl_fao<-read_sf(dsn=file.path(dir_B,"prep/spatial/output/Regions_atl_FAO.shp"),layer="Regions_atl_FAO") %>% 
  filter(F_LEVEL=="MAJOR")
rgns_lng_atl<-read_sf(dsn=file.path(dir_B,"prep/spatial/output/atl_ohi_longhurst.shp"),layer="atl_ohi_longhurst")
rgns_lng_atl<-mutate(rgns_lng_atl, id_lng=rownames(rgns_lng_atl))
fao_lng<-st_intersection(rgns_atl_fao,rgns_lng_atl)

lng_fao<-st_intersection(rgns_lng_atl,rgns_atl_fao)
plot(lng_fao["id_lng"])

crs(rgns_atl_fao)
crs(rgns_lng_atl)
dev.neW()
plot(fao_lng["id_lng"],col=brewer.pal(20, "Set1"))#see R tips and tricks for colour
plot(fao_lng["F_AREA"])
writeOGR(fao_lng,dsn=file.path(dir_B, "prep/spatial/output"), layer = "fao_lng", driver="ESRI Shapefile")

st_write(fao_lng,dsn='~/big/prep/spatial/output',layer='fao_lng', driver = "ESRI Shapefile")

fao_lng<-as(fao_lng,"spatial")
fao_lng$id_lng
```

```{r}

#raster of the atlantic FAO regions created with zonal in npp_pres layer
zones <- stack(file.path(bigdir, 'zones.tif'))#Large SpatialGridDataFrame


library(ncdfgeom)
library(maptools)
data(wrld_simpl)
plot(wrld_simpl)

```
##Longhurst provinces
```{r}
lng_gl<-st_read("~/big/Longhurst_world_v4_2010/Longhurst_world_v4_2010.shp",quiet=TRUE)#I have simplified the geometry through the mapshaper.org. The problem lies in some polygons of the baltic sea.
#
lng_gl_sp <- as(lng_gl, 'Spatial')

#lng_gl_sptr <- spTransform(lng_gl_sp, CRS( "+init=epsg:4326" ) ) 
lng_gl_sp_buffer <- rgeos::gBuffer(lng_gl_sp, byid=TRUE, width=0) # go in to solve topology problems


head(lng_gl)
crs(lng_gl_sp)
plot(lng_gl_sp)
head(lng_gl_sptr)
lng_atl_cr<-raster::crop(lng_gl_sp_buffer,reg_atl_FAO_sp)
#lng_atl<-raster::mask(lng_atl_cr,reg_atl_FAO_sp)Error in (function (classes, fdef, mtable) : unable to find an inherited method for function ‘mask’ for signature ‘"SpatialPolygonsDataFrame", "SpatialPolygonsDataFrame"’

lng_atl_cr<-raster::crop(lng_gl_sp_buffer,atl_mask)#questa non funziona-forse non hanno la stessa proezione
rgns_lng_atl<-lng_atl_cr
plot(lng_atl_cr)
head(lng_atl_cr)



```



#Save shapefile

Use rgdal::writeOGR() to save the shapefile. Set the driver equal to ESRI Shapefile.

```{r save}
#writeOGR(lng_atl_cr,dsn='~/big/prep/spatial/output',layer='atl_ohi_longhurst', driver = "ESRI Shapefile")

#save a copy on github
writeOGR(rgns_lng_atl,dsn='C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output',layer='rgns_lng_atl', driver = "ESRI Shapefile")

rgns_lng_atl<- st_read("C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output/rgns_lng_atl.shp", quiet = TRUE)

plot(rgns_lng_atl[1])+plot(inters[1])

##add a column with number id for each provinces:

rgns_lng_atl<-mutate(rgns_lng_atl, id_lng=rownames(rgns_lng_atl))
inters<-st_intersection(rgns_lng_atl,regions_atl_FAO)
plot(inters[1])
inters_sp
plot(lng_atl_cr)
raster::click(lng_atl_cr)
```


##LME provinces

```{r}

lme_gl<-st_read("~/big/prep/spatial/raw/LMEs66.shp",quiet=TRUE)
#I have simplified the geometry through the mapshaper.org. The problem lies in some polygons of the baltic sea.
#
lme_gl_sp <- as(lme_gl, 'Spatial')

lme_gl_sp_buffer <- rgeos::gBuffer(lme_gl_sp, byid=TRUE, width=0) # go in to solve topology problems


head(lme_gl)
crs(lme_gl_sp)==crs(reg_atl_FAO_sp)
plot(lme_gl_sp)
head(lng_gl_sptr)
lme_atl_cr<-raster::crop(lme_gl_sp_buffer,reg_atl_FAO_sp)

plot(lme_atl_cr)
head(lng_atl_cr)


```

```
```{r}
```{r save}
#writeOGR(lme_atl_cr,dsn='~/big/prep/spatial/output',layer='reg_atl_lme', driver = "ESRI Shapefile")

#save a copy on github
writeOGR(lme_atl_cr,dsn='C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output',layer='reg_atl_lme', driver = "ESRI Shapefile")

reg_atl_lme <- st_read("C:/Users/szunino/Documents/github/Atl_ohi/prep/spatial/output/reg_atl_lme.shp", quiet = TRUE)

plot(reg_atl_lme[3])
```


## get a sample file to convert other spatial data
tmp <- read_sf(dsn = RAM_spatial_dir, layer = ram_sf[1])
```

## Get OHI Raster/Zonal Data:

* **regions_eez_with_fao_ant.tif** - This includes all the ocean regions (eez/fao/antarctica), but the raster cell values correspond to the rgn_ant_id in regions_2017_update.  This file is most often used to extract pressure values for each region.
* **regionData.csv** - has data for spatial id's used in raster

#Data Source: https://mazu.nceas.ucsb.edu/data/
Regions: This folder includes shapefiles in both the Mollweide (regions_mol) and WGS 84 (regions_gcs) coordinate reference systems (Fig. 1). See this primer for more information about coordinate reference systems. These files can be opened in mapping programs such as ArcGIS, GRASS GIS, and R. Table 1 includes information about the data fields included in these shapefiles.

```{r, eval=FALSE}
## Grab OHI base data for pressures
zones <- raster(file.path(dir_B, "git-annex/globalprep/spatial/v2017/regions_eez_with_fao_ant.tif")) 
rgn_data <- read.csv(file.path(dir_M, "git-annex/globalprep/spatial/d2014/data/rgn_mol_raster_1km/regionData.csv"))


```{r RAM to OHI_FAO rgn, eval = FALSE}
##################################
## Combine OHI and FAO regions
## get OHI regions and transform



regions <- read_sf(dsn = file.path(dir_B, "prep/spatial/spatial_ohi_supplement/regions"), layer = "regions_gcs") %>% 
  filter(rgn_typ!="land") 
plot(regions)
regions_atl_FAO <- st_read("~/big/prep/spatial/output/Regions_atl_FAO.shp", quiet = TRUE)
## get a sample file to convert other spatial data
tmp <- read_sf(dsn = RAM_spatial_dir, layer = ram_sf[1])

regions_atl_tr <- st_transform(regions_atl_FAO, st_crs(tmp)$epsg)
regions_tr <- st_transform(regions, st_crs(tmp)$epsg)
regions_ohi_fao_atl <- st_intersection(regions_tr, regions_atl_tr)
head(regions_ohi_fao_atl)
st_write(regions_ohi_fao_atl, dsn = file.path(dir_B, "prep/spatial/int"), 
                             layer = "regions_atl_tr", driver = "ESRI Shapefile")
```


```{r RAM to OHI_FAO rgn, eval = FALSE}
##I will add the id code of the LME and LNG areas.
reg_atl_LNG<-as(reg_atl_LNG,"spatial")
rgns_atl_tr <- st_transform(reg_atl_LNG, st_crs(tmp)$epsg)
regions_ohi_fao_lng_atl<-st_intersection(regions_ohi_fao_atl, regions_lng_tr)

st_write(regions_ohi_fao_lng_atl, dsn = file.path(dir_B, "prep/spatial/int"), 
                             layer = "regions_ohi_fao_lng_atl", driver = "ESRI Shapefile")

##regions_atl_sp <- as(regions_atl, "Spatial") # convert to spatial polygons 

## FAO region

#fao_regions <- read_sf(dsn = file.path(dir_M, #"git-annex/Global/NCEAS-Regions_v2014/orig/FAO_AREAS"), layer="FAO_AREAS") %>%
#  filter(F_LEVEL == "MAJOR")
str(regions_atl_tr)

st_write(regions_atl_tr, dsn = file.path(dir_B, "prep/spatial/int"), 
                             layer = "regions_atl_tr", driver = "ESRI Shapefile")

sf::write_sf(regions_atl,paste0(spatial_output),  "Regions_atl.shp", driver = "ESRI Shapefile",overwrite=TRUE)
############################################
## Overlay stock data on fao-ohi region data

fao_ohi <- st_read(file.path(dir_M, "git-annex/globalprep/fis/v2017/int"), 
        layer = "ohi_fao_rgns")

stock_fao_ohi <- NA
for(stock in ram_sf) {   # stock = ram_sf[85]
    cat(paste0(stock, "\n"))
    tmp_poly <- read_sf(dsn = RAM_spatial_dir, layer = stock)
    tmp_fao_ohi <- st_intersection(fao_ohi, tmp_poly)
    if(sum(is.na(stock_fao_ohi))==1){
      stock_fao_ohi <- tmp_fao_ohi
    }else
    {stock_fao_ohi <- rbind(stock_fao_ohi, tmp_fao_ohi)}
}
```


```{r RAM to OHI_FAO rgn, eval = FALSE}
##################################
from v2017 Ram data prep_I need this script to merge OHI and FAO data and eventually merge lng or the new shp that will come out from atlanteco
## Combine OHI and FAO regions
## get OHI regions and transform
ohi <- read_sf(dsn = file.path(dir_M, "git-annex/globalprep/spatial/v2017"), layer = "regions_2017_update")
ohi <- st_transform(ohi, st_crs(tmp)$epsg)
# ohi <- as(ohi, "Spatial") # convert to spatial polygons file
## FAO region
fao_regions <- read_sf(dsn = file.path(dir_M, "git-annex/Global/NCEAS-Regions_v2014/orig/FAO_AREAS"), layer="FAO_AREAS") %>%
  filter(F_LEVEL == "MAJOR")
fao_ohi <- st_intersection(ohi, fao_regions)
st_write(fao_ohi, dsn = file.path(dir_M, "git-annex/globalprep/fis/v2017/int"), 
                             layer = "ohi_fao_rgns", driver = "ESRI Shapefile")
############################################
## Overlay stock data on fao-ohi region data
fao_ohi <- st_read(file.path(dir_M, "git-annex/globalprep/fis/v2017/int"), 
        layer = "ohi_fao_rgns")
stock_fao_ohi <- NA
for(stock in ram_sf) {   # stock = ram_sf[85]
    cat(paste0(stock, "\n"))
    tmp_poly <- read_sf(dsn = RAM_spatial_dir, layer = stock)
    tmp_fao_ohi <- st_intersection(fao_ohi, tmp_poly)
    if(sum(is.na(stock_fao_ohi))==1){
      stock_fao_ohi <- tmp_fao_ohi
    }else
    {stock_fao_ohi <- rbind(stock_fao_ohi, tmp_fao_ohi)}
}