---
title: "shift long"
author: "sz"
date: "3/2/2021"
output: html_document
---

count_density_size1_360
```{r}
  weight_torepro = list.files(file.path(data_wd,'/original_eriksen_data'),'weight_*',full.names=T)
    count_torepro  = stack(list.files(file.path(data_wd,'tmp/unlog'),'unlog_*',full.names=T))
    
library(raster)
    data_wd = file.path(dir_B,'R_raster_tmp/trash')

wg <- raster(file.path(data_wd ,'original_eriksen_data/weight_density_size1_360.tif'))
wg1<- crop(wg, extent(-180, 360, -90, 90))
plot(wg1)    
wg2 <- crop(wg, extent(0, 180, -90, 90))   
plot(wg2)
    

extent(wg1) <- c(-180, 0, -90, 90)
extent(wg2)<- c(0, 180, -90, 90)

plot(wg1)

wg_merge <- merge(wg1, wg2,overlap=FALSE)

plot(wg_merge)
latlong=crs("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
crs(ma)<-"+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

plot(ma)
#writeRaster(ma, 'tmp/ma.tif', format="GTiff")
#merge<-raster("tmp/ma.tif")
#plot(merge)
??col
```


```{r}
#predno un file script e ripeto l'operaizione per 4 volte
mytheme <- rasterTheme(region = rev(colorRampPalette(brewer.pal(11, 'Spectral'))(255)))

#set working directory on neptune
data_wd = file.path(dir_B,'R_raster_tmp/trash')
setwd(data_wd)
getwd()

x <- raster(file.path(data_wd ,'original_eriksen_data/count_density_size1_360.tif'))
plot(x)
y <- raster(file.path(data_wd ,'original_eriksen_data/weight_density_size1_360.tif'))

x2 <- raster(file.path(data_wd ,'sa-marine-debris-count_density_size2_360.tif'))
x3 <- raster(file.path(data_wd ,'sa-marine-debris-count_density_size3_360.tif'))
```
```{r}
data_tria2l = file.path(dir_B,'R_raster_tmp/trash/geotiff_geoserver_debris')
prova4<-raster(file.path(data_tria2l,'ebsa-marine-debris-count_density_size4_360.tif'))
plot(prova4)
click(prova4)

```

```{r}
#Cut in two pieces, change the extent of the western hemisphere, and merge again.
x<-raster(file.path(data_wd,'original_eriksen_data/count_density_size4_360.tif'))
x1<-x
x1<- crop(x, extent(150, 360, -90, 90))
plot(x1)
```



```{r}
x2 <- crop(x, extent(0, 150, -90, 90))   
plot(x2)
```


```{r}


extent(x1) <- c(-150, 0, -90, 90)
extent(x2)<- c(0, 150, -90, 90)

ma <- merge(x1, x2,overlap=FALSE)
latlong=crs("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")
crs(ma)<-"+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"

plot(ma)
#writeRaster(ma, 'tmp/ma.tif', format="GTiff")
#merge<-raster("tmp/ma.tif")
#plot(merge)
??col
```




```{r}
extent(ma)
ocean = raster(file.path(dir_B,'prep/spatial/spatial_ohi_supplement/ocean.tif'))
plot(ocean)
crs(ma)=crs(ocean)
extent(ma)=extent(ocean)
projection(ma) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"  

writeRaster(ma,filename=file.path(data_wd,"sa-marine-debris-count_density_size4_360.tif"), format="GTiff",overwrite=TRUE)
??writeRaster
ma <- projectRaster(ma, crs=(ocean))
plot(ma)+
  plot(ocean, add=TRUE)
```


,over=T,progress='text',filename=file.path(data_wd,'tmp/prll.tif'),overwrite=T)
writeRaster(ma, 'tmp/merdall.tif',format="GTiff",overwrite=T)
extent(ma)
extent(ocean)
#This assumes a well behaved global raster with edges at -180, and 180 and the zero meridian being in between two columns.
getwd()
mollCRS=crs('+proj=moll +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs')
latlong=crs("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0")

```{r}
xyz_ch<-read.csv("Xplus180%360meno180.csv", header=TRUE)
celia_rs <- rasterFromXYZ(xyz_ch, crs=latlong)
writeRaster(celia_rs, 'tmp/celia.tif', format="GTiff")
plot(celia_rs)
cells=projectRaster(celia_rs, crs=(ocean))
plot(ocean)+
  plot(cells,add=TRUE)
crs(cells_raster)
crs(ma)
proj4string(ma)
proj4string(ocean)

extent(ma)=extent(ocean)
plot(cells)+
plot(ocean)
#metodo celia con python
new_Lon = ( old_Lon + 180 )%360 -180
```

